### 2017-08-30 model/old/overflowmcmc.tex
moved old model stuff to an old folder

\begin{document}
\section{Introduction}
This note describes the MCMC algorithm used to fit the overflow model. The bulk of this material may end up in a technical appendix or as a section of the paper. Section \ref{sec:model} briefly describes the model and section \ref{sec:mcmc} describes the MCMC strategy for computing the posterior distribution.

@@

\subsection{Transaction arrival process}
We will assume that transactions arrive according to a non-homogenous Poisson process. Let $N(\tau)$ denote the number of transactions that have arrive by time $\tau$, and let $\lambda(\tau)$ denote some intensity function such that $\lambda(\tau) > 0$ for all times $\tau$. Then any collection of non-overlapping increments $\left\{N(\tau_i + \delta_i) - N(\tau_i)\right\}$ where $\delta_i>0$ for all $i$ are independent with distribution
\begin{align*}
N(\tau_i + \delta_i) - N(\tau_i) \sim Poi(\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i))
\end{align*}
where 
\begin{align*}
\Lambda(\tau) = \int_0^\tau \lambda(u)du.
\end{align*}
Suppose the intensity function is a gamma process with shape function $\Psi(\tau)$ and rate parameter $\theta$. In other words
\begin{align*}
\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i) \stackrel{ind}{\sim} G(\Psi(\tau_i + \delta_i) - \Psi(\tau_i), \theta)
\end{align*}
with mean $[\Psi(\tau_i + \delta_i) - \Psi(\tau_i)]/\theta$. Then for block $k=1,2,\dots,n_{b}$ define $\lambda_k = \Lambda(\tau_k) - \Lambda(\tau_{k-1})$, $\eta_k = N(\tau_k) - N(\tau_{k-1})$, and $\psi_k = \Psi(\tau_k) - \Psi(\tau_{k-1})$. Then we assume
\begin{align*}
\eta_k | \lambda_{1:n_{b}} &\stackrel{ind}{\sim} Poi(\lambda_k)\\
\lambda_k|\delta_{1:n_{b}} &\stackrel{ind}{\sim} G(\psi_k, \theta).
\end{align*}

@@

We can represent this part of the model more compactly by integrating out the $\lambda_k$'s. Define $\phi = \theta/(1 + \theta)$. Then $\eta_k$ has a negative binomial distribution:
\begin{align*}
\eta_k|\delta_{1:n_{b}} &\stackrel{ind}{\sim} NB(\psi_k, 1-\phi)
\end{align*}
with pmf
\begin{align*}
P(\eta_k = \eta|\delta_{1:n_{b}}) \frac{\Gamma(\psi_k + \eta)}{\Gamma(\eta + 1)}\phi^{\psi_k}(1-\phi)^{\eta}
\end{align*}
for $\eta=0,1,\dots$ We will use this representation of the model both for simplified exposition and later for cheaper MCMC.

@@

\subsection{Prior on $\sigma^2$}
The expected size of the measurement error on $t_k$ is controlled by $\sigma^2$ --- the larger $\sigma^2$, the farther away we expect $t_k$ to be from $\tau_k$. We will use a half-$t$ prior on $\sigma$ which can be written as the following prior on $\sigma^2$:
\begin{align*}
\sigma^2|\omega &\sim IG(v/2, \omega); && \omega \sim G(1/2, s^2/2).
\end{align*}
This yields a $t_{v}(0,s^2)1(\sigma>0)$ prior on $\sigma$ where $s^2$ is a scale parameter and $v$ is a degrees of freedom parameter. We use $v=1$ so that the prior on $\sigma$ has a half-Cauchy distribution with median $s$. To gain some intuition in order to set an appropriate value for $s$, suppose that the measurement error distribution was not truncated. Then we would expect $68\%$ of observed block arrival times to be within $\sigma$ of the actual arrival time. A reasonable a priori guess for $\sigma$ is then one minute, so we set $s=1$ so that the median value of $\sigma$ in its prior is one. This prior is only weakly informative --- the right tail is gently sloped so that the data will quickly overwhelm the prior as observations pile up [CITE GELMAN VARIANCE PAPER].

@@

\subsection{Prior on $\bm{\gamma}$}
As we will see in the next section, there is no convenient conditionally conjugate form for $\bm{\gamma}$, so there is no computational trade-off associated with choosing a prior which more accurately reflects prior knowledge. We will consider $\bm{\gamma}$ much like a regression parameter and set
\begin{align*}
\bm{\beta} \sim N(\bm{g}_\gamma, \bm{S}_\gamma)
\end{align*}
where $\bm{g}_\gamma$ is a mean vector and $\bm{S}_\gamma$ is a covariance matrix. We will use a weakly informative prior with $\bm{g}_\gamma = \bm{0}$ and $\bm{S}_\gamma = s^2_\gamma\bm{I}$ where $s_\gamma = 100$.  [MAYBE SOMETHING WITH FATTER TAILS HERE? MAYBE NOTE THAT NONINFORMATIVE PRIOR IS NOT A GOOD IDEA SINCE WE WANT TO GUARANTEE POSTERIOR PROPRIETY --- SAME FOR $s_\gamma$ TOO LARGE (VIVEK'S PAPER)]

\subsection{Priors on $\alpha$ and $\beta$}
The conditionally conjugate prior for $beta$ is a gamma while $\alpha$ has no convenient conditionally conjugate form. We will assume that $\beta \sim G(a_\beta, b_\beta)$ and $\alpha \sim G(a_\alpha, b_\alpha)$. In order to choose $a_\alpha$, $b_\alpha$, $a_\beta$, and $b_\beta$, consider that $E[d_i|\alpha,\beta] = \alpha/\beta$. But 
\begin{align*}
E\left[\frac{\alpha}{\beta}\right] = E[\alpha]E\left[\frac{1}{\beta}\right] = \frac{a_\alpha}{b_\alpha}\frac{b_\beta}{a_\beta - 1}.
\end{align*}
Similarly
\begin{align*}
V\left[\frac{\alpha}{\beta}\right] = V[\alpha] + V\left[\frac{1}{\beta}\right] = \frac{a_\alpha}{b_\alpha^2} + \frac{b_\beta^2}{(a_\beta - 1)^2(a_\beta - 2)}.
\end{align*}
We will choose these parameters so that $E[\alpha/\beta] \approx 0.01$ MB and $V[\alpha/\beta]$ is large so that the prior is weakly informative. We will set $a_\beta = 2.1$ to ensure the variance calculated above exists. Then $a_\alpha = 0.1$, $b_\alpha = 1$, and $b_\beta = 1$ yields
\begin{align*}
E\left[\frac{\alpha}{\beta}\right] = \frac{1}{21} \approx 0.09; && V\left[\frac{\alpha}{\beta}\right] \approx 10.
\end{align*}

@@

In order to simulate from the posterior we construct a Gibbs sampler using data augmentation to expand the state space and Metropolis steps where appropriate. We use data augmentation in two ways. First, in order to more effectively deal with the truncated distribution in the measurement error model for block arrival times we introduce $\tilde{\bm{t}}_k=(\tilde{t}_{1k},\dots,\tilde{t}_{\tilde{n}_kk}, t_k)$ as the original $t_k$ along with $\tilde{n}_k$ draws from the untruncated normal distribution for $t_k$ that occurred before $t_k$ was drawn from the restricted support. Second, the models for $\eta_{1:n_{b}}$ and $d_{1:n_d}$ induce a model for $(M_{1:n_{b}},D_{1:n_d})$ since the latter are a detrministic function of the former, but this model is not easy work with. Instead, we explicitly use $\eta_{1:n_{b}}$ and $d_{1:n_d}$ in the Gibbs sampler as additional data augmentation.

\subsection{Data augmentation for truncated distributions --- $\tilde{n}_k$ and $\tilde{t}_k$}
For block $k=1,2,n_{b}$ the measurement error model on block arrival times is
\begin{align*}
t_k|t_{1:(k-1)} \sim N(\tau_k,\sigma^2)1(t_k > m_k)
\end{align*}
where $m_k = \mathrm{median}(t_{k - 1:11})$ with density 
\begin{align*}
p(\bm{t_{1:n_{b}}}|\bm{\tau}_{1:n_{b}},\sigma^2) \propto \sigma^{-n_{b}}\frac{\exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}(t_k - \tau_k)^2\right]}{\prod_{k=1}^{n_{b}}\left[1 - \Phi\left(\frac{m_k - \tau_k}{\sigma}\right)\right]}
\end{align*}
where $\Phi(.)$ is the standard normal cdf. Drawing from the full conditional distributions of $\sigma^2$ and $\tau_k$ can be challenging since both parameters enter into the normalizing constant of $p(\bm{t_{1:n_{b}}}|\bm{\tau}_{1:n_{b}},\sigma^2)$. Consider an underlying process which gives rise to the truncated distribution --- instead of simply drawing $t_k$ from the truncated normal distribution, this process draws from the untruncated normal distribution until it obtains a draw from the restricted region of the parameter space, i.e. until it successfully obtains a $t_k > m_k$. Let $\tilde{n}_k$ be the number of failures before the observed success $t_k$. Then $\tilde{n}_k$ has a negative binomial distribution, specifically $\tilde{n}_k \sim NB(1, \rho_k)$ with density
\begin{align*}
p(\tilde{n}_k|\tau_k,\sigma^2) = \rho_k^{\tilde{n}_k}(1 - \rho_k)
\end{align*}
where $\rho_k$ is the probability of failure, i.e. $\rho_k = \Phi([m_k - \tau_k]/\sigma)$. Conditional on $\tilde{n}_k$ the density of the $\tilde{n}_k$ failures, $\tilde{t}_{1k},\dots,\tilde{t}_{\tilde{n}_kk}$, is again truncated normal, but this time truncated to the opposite end of the parameter space. That is for $i = 1,\dots,\tilde{n}_k$
\begin{align*}
\tilde{t}_{ik}\stackrel{iid}{\sim}N(\tau_k,\sigma^2)1(t_k < m_k).
\end{align*}
So the data augmentation step of the Gibbs sampler consists of the following substeps: for $k=1,2,\dots,n_{b}$,
\begin{enumerate}
\item Draw $\tilde{n}_k \sim NB(1,\rho_k)$ where $\rho_k = \Phi([ m_k -  \tau_k]/\sigma)$.
\item If $\tilde{n}_k > 0$, for $i=1,\dots,\tilde{n}_k$ draw $\tilde{t}_{ik} \sim N(\tau_k,\sigma^2)1(t_k < m_k)$ and form $\tilde{\bm{t}}_k = (\tilde{t}_{1k},\dots,\tilde{t}_{\tilde{n}_kk}, t_k)'$.
\end{enumerate}
For simplicity in the other Gibbs steps define $\tilde{t}_{\tilde{n}_k + 1,k}\equiv t_k$.

It is tempting to imbue $\tilde{n}_k$ and $\tilde{t_{ik}}$ with some interpretation relative to the model, e.g. as in [CITE GELMAN PAPER]. For example, $\tilde{n}_k$ may represent the number of blocks that were rejected by the Bitcoin protocol for having a timestamp earlier than the median of the previous elevent blocks. While tempting this interpretation is not strictly speaking correct. When a block is rejected by the protocol a new block does not appear instantaneously since it takes time for another miner to discover the block. So the arrival times and thus measurement error distributions of the rejected block and the accepted block are not the same. While the interpretation is not correct, it is good to be inspired by computational tricks to create better model [CITE SAME GELMAN PAPER] and perhaps something similar to this negative binomial structure can be used to add rejected blocks as a component of the model, though we do not explore this possibility here.

\subsection{Measurement error variance --- $\sigma^2$}
The full conditional distribution of the measurement error variance, $\sigma^2$, depends only on $\tilde{\bm{t}}_{1:n_{b}}$ and $\tau_{1:n_{b}}$. It is
\begin{align*}
p(\sigma^2|...) \propto \sigma^{-(n_{b} + \sum_{k=1}^{n_{b}}\tilde{n}_k)}\exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k}(\tilde{t}_{ik} - \tau_k)^2\right]p(\sigma^2)
\end{align*}
where $p(\sigma^2)$ is the density of the prior distribution on $\sigma^2$. With the half-$t$ prior on $\sigma$ discussed above and letting $a_\sigma = (n_{b} + \sum_{k=1}^{n_b}\tilde{n}_k)/2$ and $b_\sigma = \sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k}(\tilde{t}_{ik} - \tau_k)^2$ the full conditional distribution of $(\sigma^2,\omega)$ is
\begin{align*}
p(\sigma^2,\omega|...) \propto (\sigma^2)^{-(a_\sigma+v)/2-1}\exp\left[-\frac{b_\sigma}{\sigma^2}\right]\exp\left[-\frac{\omega}{\sigma^2}\right]\omega^{(v + 1)/2 - 1}\exp\left[-\frac{\omega}{v s^2}\right].
\end{align*}
So we draw $\sigma^2$ and $\omega$ in two separate Gibbs steps:
\begin{enumerate}
\item Draw $\omega \sim G\left(\frac{v + 1}{2}, \frac{1}{\sigma^2} + \frac{1}{v s^2}\right)$ and form\footnote{The inconsistency in notation between the definition of $a_\sigma$ and $b_\sigma$ is a consequence of matching notation to the R code supplied for implementing this MCMC algorithm.}
\begin{align*}
a_\sigma = \frac{n_{b} + \sum_{k=1}^{n_b}\tilde{n}_k + v}{2} &&\mbox{  and  }&& b_\sigma = \sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k}(\tilde{t}_{ik} - \tau_k)^2;
\end{align*}
\item Draw $\sigma^2 \sim  IG\left(a_\sigma, b_\sigma/2 + \omega\right)$.
\end{enumerate}

\subsection{Block arrival times --- $\tau_k$}
The full conditional distribution of the block arrival times is more complicated:
\begin{align*}
p(\bm{\tau}_{1:n_{b}}|...) \propto & \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \tau_k\right)^2 - \frac{1}{10}\tau_{n_{b}}\right]\\
&\times \prod_{k=1}^{n_b}\Gamma[\psi_k(\tau_{k}, \tau_{k-1}) + \eta_k]\phi^{\psi_k(\tau_k,\tau_{k-1})}1(0\le \tau_1 \le \tau_2 \le \hdots \le \tau_{n_b})
\end{align*}
where $\tau_0\equiv0$ and $\Gamma(x) = \int_0^\infty t^{x-1}e^{-t}dt$ is the gamma function. Here we make explicit that $\psi_k$ is a function of $\tau_k$ and $\tau_{k-1}$. Even if $\Psi(\tau)$ is some nice known form, this density is nonstandard because the $\tau_k$'s enter the gamma function. Some ideas for handling this step:
\begin{enumerate}
\item Draw each $\tau_k$ condtional on all the others using random walk Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_k$ sufficiently well and use independent Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_{1:n_{b}}$ sufficiently well and use an independent Metropolis step.
\item Use adaptive rejection sampling (ARS) to draw from the density of $\tau_k$.
\item Use ARMS --- ARS except with a Metropolis correction for densities which are not log concave (look it up).
\end{enumerate}

Let $\delta_k = \tau_k - \tau_{k-1}$ for $k=1,2,\dots,n_b$. Then the determinant of the Jacobian is $1$ full conditional of $\delta_{1:n_b}$ is
\begin{align*}
p(\bm{\delta}_{1:n_{b}}|...) \propto & \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \sum_{l=1}^k\delta_l\right)^2 - \frac{1}{10}\sum_{k=1}^{n_b}\delta_{k}\right]\\
&\times \prod_{k=1}^{n_b}\Gamma(\psi_k + \eta_k)\phi^{\psi_k}1(0\le \delta_k).
\end{align*}
Finally let $\zeta_k = \log(\delta_k)$ for $k = 1, 2, \dots, n_b$. Then the determinate of the Jacobian is $\prod_{k=1}^{n_b}e^{\zeta_k}$ and the full conditional of $\zeta_{1:n_b}$ is
\begin{align*}
p(\bm{\zeta}_{1:n_{b}}|...) \propto & \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \sum_{l=1}^ke^{\zeta_l}\right)^2 - \frac{1}{10}\sum_{k=1}^{n_b}e^{\zeta_k} + \sum_{k=1}^{n_b}{\zeta_k}\right] \prod_{k=1}^{n_b}\Gamma(\psi_k + \eta_k)\phi^{\psi_k}.
\end{align*}

>>>>

\begin{document}
\section{Introduction}
This note describes the MCMC algorithm used to fit the overflow model. The bulk of this material may end up in a technical appendix or as a section of the paper. Section \ref{sec:model} briefly describes the model and Section \ref{sec:mcmc} describes the MCMC strategy for computing the posterior distribution.

@@

\subsection{Transaction arrival process}
We will assume that transactions arrive according to a non-homogenous Poisson process. Let $N(\tau)$ denote the number of transactions that have arrived by time $\tau$, and let $\lambda(\tau)$ denote some intensity function such that $\lambda(\tau) > 0$ for all times $\tau$. Then any collection of non-overlapping increments $\left\{N(\tau_i + \delta_i) - N(\tau_i)\right\}$ where $\delta_i>0$ for all $i$ are independent with distribution
\begin{align*}
N(\tau_i + \delta_i) - N(\tau_i) \sim Poi(\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i))
\end{align*}
where 
\begin{align*}
\Lambda(\tau) = \int_0^\tau \lambda(u)du.
\end{align*}
Suppose the intensity function is a gamma process with shape function $\Psi(\tau)$ and rate parameter $\theta$. In other words
\begin{align*}
\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i) \stackrel{ind}{\sim} G(\Psi(\tau_i + \delta_i) - \Psi(\tau_i), \theta)
\end{align*}
with mean $[\Psi(\tau_i + \delta_i) - \Psi(\tau_i)]/\theta$. Then for block $k=1,2,\dots,n_{b}$ define $\lambda_k = \Lambda(\tau_k) - \Lambda(\tau_{k-1})$, $\eta_k = N(\tau_k) - N(\tau_{k-1})$, and $\psi_k = \Psi(\tau_k) - \Psi(\tau_{k-1})$. Then we assume
\begin{align*}
\eta_k | \lambda_{1:n_{b}} &\stackrel{ind}{\sim} Poi(\lambda_k)\\
\lambda_k|\delta_{1:n_{b}} &\stackrel{ind}{\sim} G(\psi_k, \theta).
\end{align*}

@@

We can represent this part of the model more compactly by integrating out the $\lambda_k$'s. Define $\phi = \theta/(1 + \theta)$. Then $\eta_k$ has a negative binomial distribution:
\begin{align*}
\eta_k|\delta_{1:n_{b}} &\stackrel{ind}{\sim} NB(\psi_k, 1-\phi)
\end{align*}
with pmf
\begin{align*}
P(\eta_k = \eta|\delta_{1:n_{b}}) = \frac{\Gamma(\psi_k + \eta)}{\Gamma(\psi_k)\Gamma(\eta + 1)}\phi^{\psi_k}(1-\phi)^{\eta}
\end{align*}
for $\eta=0,1,\dots$ We will use this representation of the model both for simplified exposition and later for cheaper MCMC.

@@

\subsection{Prior on $\sigma^2$}
The expected size of the measurement error on $t_k$ is controlled by $\sigma^2$ --- the larger $\sigma^2$, the farther away we expect $t_k$ to be from $\tau_k$. We will use a half-$t$ prior on $\sigma$ which can be written as the following prior on $\sigma^2$:
\begin{align*}
\sigma^2|\omega &\sim IG(v/2, \omega); && \omega \sim G(1/2, 1/\{2vs^2\}).
\end{align*}
This yields a $t_{v}(0,s^2)1(\sigma>0)$ prior on $\sigma$ where $s^2$ is a scale parameter and $v$ is a degrees of freedom parameter. We use $v=1$ so that the prior on $\sigma$ has a half-Cauchy distribution with median $s$. To gain some intuition in order to set an appropriate value for $s$, suppose that the measurement error distribution was not truncated. Then we would expect $68\%$ of observed block arrival times to be within $\sigma$ of the actual arrival time. A reasonable a priori guess for $\sigma$ is then one minute, so we set $s=1$ so that the median value of $\sigma$ in its prior is one. This prior is only weakly informative --- the right tail is gently sloped so that the data will quickly overwhelm the prior as observations pile up [CITE GELMAN VARIANCE PAPER].

@@

\subsection{Prior on $\bm{\gamma}$}
As we will see in the next section, there is no convenient conditionally conjugate form for $\bm{\gamma}$, so there is no computational trade-off associated with choosing a prior which more accurately reflects prior knowledge. We will consider $\bm{\gamma}$ much like a regression parameter and set
\begin{align*}
\bm{\gamma} \sim N(\bm{g}_\gamma, \bm{S}_\gamma)
\end{align*}
where $\bm{g}_\gamma$ is a mean vector and $\bm{S}_\gamma$ is a covariance matrix. We will use a weakly informative prior with $\bm{g}_\gamma = \bm{0}$ and $\bm{S}_\gamma = s^2_\gamma\bm{I}$ where $s_\gamma = 100$.  [MAYBE SOMETHING WITH FATTER TAILS HERE? MAYBE NOTE THAT NONINFORMATIVE PRIOR IS NOT A GOOD IDEA SINCE WE WANT TO GUARANTEE POSTERIOR PROPRIETY --- SAME FOR $s_\gamma$ TOO LARGE (VIVEK'S PAPER)]

\subsection{Priors on $\alpha$ and $\beta$}
The conditionally conjugate prior for $\beta$ is a gamma while $\alpha$ has no convenient conditionally conjugate form. We will assume that $\beta \sim G(a_\beta, b_\beta)$ and $\alpha \sim G(a_\alpha, b_\alpha)$. In order to choose $a_\alpha$, $b_\alpha$, $a_\beta$, and $b_\beta$, consider that $E[d_i|\alpha,\beta] = \alpha/\beta$. But 
\begin{align*}
E\left[\frac{\alpha}{\beta}\right] = E[\alpha]E\left[\frac{1}{\beta}\right] = \frac{a_\alpha}{b_\alpha}\frac{b_\beta}{a_\beta - 1}.
\end{align*}
Similarly
\begin{align*}
V\left[\frac{\alpha}{\beta}\right] = V[\alpha] + V\left[\frac{1}{\beta}\right] = \frac{a_\alpha}{b_\alpha^2} + \frac{b_\beta^2}{(a_\beta - 1)^2(a_\beta - 2)}.
\end{align*}
We will choose these parameters so that $E[\alpha/\beta] \approx 0.01$ MB and $V[\alpha/\beta]$ is large so that the prior is weakly informative. We will set $a_\beta = 2.1$ to ensure the variance calculated above exists. Then $a_\alpha = 0.1$, $b_\alpha = 1$, and $b_\beta = 1$ yields
\begin{align*}
E\left[\frac{\alpha}{\beta}\right] = \frac{1}{21} \approx 0.09; && V\left[\frac{\alpha}{\beta}\right] \approx 10.
\end{align*}

@@

In order to simulate from the posterior we construct a Gibbs sampler using data augmentation to expand the state space and Metropolis steps where appropriate. We use data augmentation in two ways. First, in order to more effectively deal with the truncated distribution in the measurement error model for block arrival times we introduce $\widetilde{\bm{t}}_k=(\widetilde{t}_{1k},\dots,\widetilde{t}_{\widetilde{n}_kk}, t_k)$ as the original $t_k$ along with $\widetilde{n}_k$ draws from the untruncated normal distribution for $t_k$ that occurred before $t_k$ was drawn from the restricted support. Second, the models for $\eta_{1:n_{b}}$ and $d_{1:n_d}$ induce a model for $(M_{1:n_{b}},D_{1:n_d})$ since the latter are a deterministic function of the former, but this model is not easy work with. Instead, we explicitly use $\eta_{1:n_{b}}$ and $d_{1:n_d}$ in the Gibbs sampler as additional data augmentation.

\subsection{Data augmentation for truncated distributions --- $\widetilde{n}_k$ and $\widetilde{t}_k$}
For block $k=1,2,n_{b}$ the measurement error model on block arrival times is
\begin{align*}
t_k|t_{1:(k-1)} \sim N(\tau_k,\sigma^2)1(t_k > m_k)
\end{align*}
where $m_k = \mathrm{median}(t_{k - 1:11})$ with density 
\begin{align*}
p(\bm{t_{1:n_{b}}}|\bm{\tau}_{1:n_{b}},\sigma^2) \propto \sigma^{-n_{b}}\frac{\exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}(t_k - \tau_k)^2\right]}{\prod_{k=1}^{n_{b}}\left[1 - \Phi\left(\frac{m_k - \tau_k}{\sigma}\right)\right]}
\end{align*}
where $\Phi(.)$ is the standard normal cdf. Drawing from the full conditional distributions of $\sigma^2$ and $\tau_k$ can be challenging since both parameters enter into the normalizing constant of $p(\bm{t_{1:n_{b}}}|\bm{\tau}_{1:n_{b}},\sigma^2)$. Consider an underlying process which gives rise to the truncated distribution --- instead of simply drawing $t_k$ from the truncated normal distribution, this process draws from the untruncated normal distribution until it obtains a draw from the restricted region of the parameter space, i.e. until it successfully obtains a $t_k > m_k$. Let $\widetilde{n}_k$ be the number of failures before the observed success $t_k$. Then $\widetilde{n}_k$ has a negative binomial distribution, specifically $\widetilde{n}_k \sim NB(1, \rho_k)$ with density
\begin{align*}
p(\widetilde{n}_k|\tau_k,\sigma^2) = \rho_k^{\widetilde{n}_k}(1 - \rho_k)
\end{align*}
where $\rho_k$ is the probability of failure, i.e. $\rho_k = \Phi([m_k - \tau_k]/\sigma)$. Conditional on $\widetilde{n}_k$ the density of the $\widetilde{n}_k$ failures, $\widetilde{t}_{1k},\dots,\widetilde{t}_{\widetilde{n}_kk}$, is again truncated normal, but this time truncated to the opposite end of the parameter space. That is for $i = 1,\dots,\widetilde{n}_k$
\begin{align*}
\widetilde{t}_{ik}\stackrel{iid}{\sim}N(\tau_k,\sigma^2)1(t_k < m_k).
\end{align*}
So the data augmentation step of the Gibbs sampler consists of the following substeps: for $k=1,2,\dots,n_{b}$,
\begin{enumerate}
\item Draw $\widetilde{n}_k \sim NB(1,\rho_k)$ where $\rho_k = \Phi([ m_k -  \tau_k]/\sigma)$.
\item If $\widetilde{n}_k > 0$, for $i=1,\dots,\widetilde{n}_k$ draw $\widetilde{t}_{ik} \sim N(\tau_k,\sigma^2)1(t_k < m_k)$ and form $\widetilde{\bm{t}}_k = (\widetilde{t}_{1k},\dots,\widetilde{t}_{\widetilde{n}_kk}, t_k)'$.
\end{enumerate}
For simplicity in the other Gibbs steps define $\widetilde{t}_{\widetilde{n}_k + 1,k}\equiv t_k$.

It is tempting to imbue $\widetilde{n}_k$ and $\widetilde{t}_{ik}$ with some interpretation relative to the model, e.g. as in [CITE GELMAN PAPER]. For example, $\widetilde{n}_k$ may represent the number of blocks that were rejected by the Bitcoin protocol for having a timestamp earlier than the median of the previous elevent blocks. While tempting this interpretation is not strictly speaking correct. When a block is rejected by the protocol a new block does not appear instantaneously since it takes time for another miner to discover the block. So the arrival times and thus measurement error distributions of the rejected block and the accepted block are not the same. While the interpretation is not correct, it is good to be inspired by computational tricks to create better model [CITE SAME GELMAN PAPER] and perhaps something similar to this negative binomial structure can be used to add rejected blocks as a component of the model, though we do not explore this possibility here.

\subsection{Measurement error variance --- $\sigma^2$}
The full conditional distribution of the measurement error variance, $\sigma^2$, depends only on $\widetilde{\bm{t}}_{1:n_{b}}$ and $\tau_{1:n_{b}}$. It is
\begin{align*}
p(\sigma^2|...) \propto \sigma^{-(n_{b} + \sum_{k=1}^{n_{b}}\widetilde{n}_k)}\exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\widetilde{n}_k}(\widetilde{t}_{ik} - \tau_k)^2\right]p(\sigma^2)
\end{align*}
where $p(\sigma^2)$ is the density of the prior distribution on $\sigma^2$. With the half-$t$ prior on $\sigma$ discussed above and letting $a_\sigma = (n_{b} + \sum_{k=1}^{n_b}\widetilde{n}_k)$ and $b_\sigma = \sum_{k=1}^{n_{b}}\sum_{i=1}^{\widetilde{n}_k}(\widetilde{t}_{ik} - \tau_k)^2$ the full conditional distribution of $(\sigma^2,\omega)$ is
\begin{align*}
p(\sigma^2,\omega|...) \propto (\sigma^2)^{-(a_\sigma+v)/2-1}\exp\left[-\frac{b_\sigma}{\sigma^2}\right]\exp\left[-\frac{\omega}{\sigma^2}\right]\omega^{(v + 1)/2 - 1}\exp\left[-\frac{\omega}{2 v s^2}\right].
\end{align*}
So we draw $\sigma^2$ and $\omega$ in two separate Gibbs steps:
\begin{enumerate}
\item Draw $\omega \sim G\left(\frac{v + 1}{2}, \frac{1}{\sigma^2} + \frac{1}{v s^2}\right)$ and form
\begin{align*}
a_\sigma = v + n_{b} + \sum_{k=1}^{n_b}\widetilde{n}_k &&\mbox{  and  }&& b_\sigma = \sum_{k=1}^{n_{b}}\sum_{i=1}^{\widetilde{n}_k}(\widetilde{t}_{ik} - \tau_k)^2;
\end{align*}
\item Draw $\sigma^2 \sim  IG\left(a_\sigma/2, b_\sigma/2 + \omega\right)$.
\end{enumerate}

\subsection{Block arrival times --- $\tau_k$}
The full conditional distribution of the block arrival times is more complicated:
\begin{align*}
p(\bm{\tau}_{1:n_{b}}|...) \propto & \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\widetilde{n}_k+1}\left(\widetilde{t}_{ik} - \tau_k\right)^2 - \frac{1}{10}\tau_{n_{b}}\right]\\
&\times \prod_{k=1}^{n_b}\Gamma[\psi_k(\tau_{k}, \tau_{k-1}) + \eta_k]\phi^{\psi_k(\tau_k,\tau_{k-1})}1(0\le \tau_1 \le \tau_2 \le \hdots \le \tau_{n_b})
\end{align*}
where $\tau_0\equiv0$ and $\Gamma(x) = \int_0^\infty t^{x-1}e^{-t}dt$ is the gamma function. Here we make explicit that $\psi_k$ is a function of $\tau_k$ and $\tau_{k-1}$. Even if $\Psi(\tau)$ is some nice known form, this density is nonstandard because the $\tau_k$'s enter the gamma function. Some ideas for handling this step:
\begin{enumerate}
\item Draw each $\tau_k$ condtional on all the others using random walk Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_k$ sufficiently well and use independent Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_{1:n_{b}}$ sufficiently well and use an independent Metropolis step.
\item Use adaptive rejection sampling (ARS) to draw from the density of $\tau_k$.
\item Use ARMS --- ARS except with a Metropolis correction for densities which are not log concave (look it up).
\end{enumerate}

Let $\delta_k = \tau_k - \tau_{k-1}$ for $k=1,2,\dots,n_b$. Then the determinant of the Jacobian is $1$ full conditional of $\delta_{1:n_b}$ is
\begin{align*}
p(\bm{\delta}_{1:n_{b}}|...) \propto & \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\widetilde{n}_k+1}\left(\widetilde{t}_{ik} - \sum_{l=1}^k\delta_l\right)^2 - \frac{1}{10}\sum_{k=1}^{n_b}\delta_{k}\right]\\
&\times \prod_{k=1}^{n_b}\Gamma(\psi_k + \eta_k)\phi^{\psi_k}1(0\le \delta_k).
\end{align*}
Finally let $\zeta_k = \log(\delta_k)$ for $k = 1, 2, \dots, n_b$. Then the determinate of the Jacobian is $\prod_{k=1}^{n_b}e^{\zeta_k}$ and the full conditional of $\zeta_{1:n_b}$ is
\begin{align*}
p(\bm{\zeta}_{1:n_{b}}|...) \propto & \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\widetilde{n}_k+1}\left(\widetilde{t}_{ik} - \sum_{l=1}^ke^{\zeta_l}\right)^2 - \frac{1}{10}\sum_{k=1}^{n_b}e^{\zeta_k} + \sum_{k=1}^{n_b}{\zeta_k}\right] \prod_{k=1}^{n_b}\Gamma(\psi_k + \eta_k)\phi^{\psi_k}.
\end{align*}
